---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
bibliography: vignettes/references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# validateHOT 🎯

<!-- badges: start -->

[![R-CMD-check](https://github.com/JoshSchramm94/validateHOT/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JoshSchramm94/validateHOT/actions/workflows/R-CMD-check.yaml)

[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)

<!-- badges: end -->

validateHOT is a package for conjoint and MaxDiff users. It provides
functions to evaluate validation tasks, run market simulations, and
convert raw utility estimates into scores that are easier to interpret.
All three components are key functions for preference measurement
techniques such as choice-based conjoint (CBC), adaptive choice-based
conjoint (ACBC), or Maximum Difference Scaling (MaxDiff). This package
is particularly relevant for the [Sawtooth
Software](https://sawtoothsoftware.com/) community who would like to
report their analysis in *R* for open science purposes. In addition, it
is compatible with other packages, for example, the ChoiceModelR package
[@ChoiceModelR]. Further, the validateHOT package is valuable for
practitioners, who would like to conduct the analyses using open-source
software.

Researchers and practitioners use preference measurement techniques for
various purposes, such as calculating the importance of specific
attributes or simulating markets [@gilbride2008; @steiner2018]. The
ultimate goal is to predict future behavior [@green1990]. To ensure
valid and reliable results, it is crucial that the collected data is
valid and can predict outcomes for tasks that were not included in the
estimation of the utility scores. Including validation tasks is highly
recommended [@Orme2015; @rao2014]. They do not only verify data validity
but can also help to test different models. The validateHOT package
offers helpful tools, to facilitate these functions (i.e., validate a
validation task, run market simulations, and communicate results of
preference measurement techniques) - all within an open-source tool.

> The validateHOT package was primarily developed for use with Sawtooth
> Software [@sawtooth2024] and the ChoiceModelR package [@ChoiceModelR]. Please be cautious about using
> it with other platforms (especially for linear and piecewise-coded
> variables).

👉🏾 <u>**What you need to provide**</u>: <br> After collecting your data
and running your initial hierarchical Bayes model, the validateHOT
package allows you to test how well your model predicts choices in the
validation task, perform market simulations or convert your data into
scores that are easier to interpret. To use the validateHOT package, you
need to import in your raw utility scores. If you plan to validate a
validation task you also need to provide the actual choice made in this
task. We provide a short tutorial in this markdown. For a more
comprehensive tutorial, please see the vignette provided with the
validateHOT package
(`vignette("validateHOT", package = "validateHOT")`).

👈🏾 <u>**What you get**</u>:<br> The validateHOT package currently
provides functions for four key components:

-   validation metrics

-   metrics commonly reported in machine learning (i.e., confusion
    matrix)

-   simulation methods, such as determining optimal product combinations

-   converting raw logit utilities into more interpretable scores

For the first three components, the `create_hot()` function is essential.
This function calculates the total utilities for each alternative in the
validation task and the market simulation.
The `create_hot()` finction computes the total utility of each alternative
based on the additive utility model [@rao2014, p. 82].

### Classical validation metrics

-   `hitrate()`: calculates the hit rate (correctly predicted choices) of
    the validation task.
-   `kl()`: Kullback-Leibler-Divergence calculates the divergence between
    the actual choice distribution and the predicted choice distribution
    [@ding2011; @philentropy]. Due to the asymmetry of the Kullback-Leibler divergence, the output includes divergence both from predicted to observed and from observed to predicted. The validateHOT
    package currently provides two logarithm bases: $log$ and $log_2$.
-   `mae()`: calculates the mean absolute error, i.e., the deviation between
    predicted and stated choice shares
-   `medae()`: calculates the median absolute error
-   `mhp()`: calculates the averaged hit probability of participant's
    actual choice in the validation task
-   `rmse()`: provides the root mean square error of deviation between
    predicted and stated choice shares

All functions can be extended with the `group` argument to get output
split by group(s).

### Confusion Matrix

The validateHOT package includes metrics from machine learning, i.e., the confusion
matrix (e.g., @burger2018). For all of the five provided functions, a
**none** alternative has to be included in the validation task. The
logic of the implemented confusion matrix is to test, for example,
whether a buy or no-buy was correctly predicted. Information could be
used to get a sense of overestimation and underestimation of general
product demand. In the table below `TP` stands for true positives, `FP`
for false positives, `TN` for true negatives, and `FN` for false
negatives [@burger2018; @kuhn2008]. To translate this to the logic of
the validateHOT package, imagine you have a validation task with five
alternatives plus the alternative of not buying.
The validateHOT package now calculates whether or not a buy (participant
opts for one of the five alternatives) or a no-buy (participant opts for
the none alternative), respectively, is correctly predicted.

Please be aware that the validateHOT package applies the following
coding of the *buy* and *no-buy* alternatives. Rows refer to the
observed decisions while columns refer to the predicted ones.

|        | Buy | No-buy |
|--------|:---:|:------:|
| Buy    | TP  |   FN   |
| No-buy | FP  |   TN   |

-   `accuracy()`: calculates the number of correctly predicted choices
    (buy or no-buy); $\frac{TP + TN}{TP + TN + FP + FN}$ [@burger2018]

-   `f1()`: defined as
    $\frac{2 * precision * recall}{precision + recall}$ or stated
    differently by @burger2018 $\frac{2TP}{2TP + FP + FN}$

-   `precision()`: defined as $\frac{TP}{TP + FP}$ [@burger2018]

-   `recall()`: defined as $\frac{TP}{TP + FN}$ [@burger2018]

-   `specificity()`: defined as $\frac{TN}{TN + FP}$ [@burger2018]

Again, all functions can be extended with the `group` argument to get
the output split by group(s).

### Simulation Methods

-   `turf()`: **T**(otal) **U**(nduplicated) **R**(each) and
    **F**(requency) is a "product line extension model"
    [@miaoulis1990; p. 29] that helps to find the perfect product bundle
    based on the reach (e.g., how many participants consider buying at
    least one product of that assortment) and the frequency (how many
    products on average are a purchase option). `turf()` currently
    provides both the *threshold* approach (`approach ='thres'`; all
    products that exceed a threshold are considered, e.g., as purchase
    option; @chrzan2019, p. 112) and the *first choice* approach
    (`approach = 'fc'`; only product with highest utility is considered
    as purchase option; @chrzan2019, p. 111).
-   `turf_ladder()`: starts with one item and then subsequently adds one
    new item, which adds the maximum possible reach. Alternatively, it
    is also possible to define fixed alternatives (i.e., alternatives
    that must be part of the assortment).
-   `freqassort()`: Similar to `turf()`, `freqassort()` will give you
    the average frequency, representing how many products the participants will
    choose from a potential assortment. Again, you have to define a
    `none` alternative, because `freqassort()` uses the *threshold*
    approach, meaning that if the utility of one product is above the utility
    of `none`, it is marked as a potential purchase option
    [@chrzan2019, p. 112]. While `turf()` calculates the reach and
    frequency for **all** combinations, you specify the combination you
    are interested in `freqassort()`.
-   `reach()`: Similar to `turf()`, `reach()` will give you the average
    percentage of how many participants you can reach (at least one of
    the products resembles a purchase option) with your in the function 
    determined potential assortment. `reach()` also uses the *threshold*
    approach (see above). While `turf()` calculates the reach and
    frequency for **all** combinations, you specify the combination you
    are interested in `reach()`.
-   `marksim()`: Runs market simulations (either the share of preference,
    `sop` or first choice rule, `fc`), including the standard error, and
    the lower and upper confidence intervals, which are calculated
    according to the following formula
    $mean +/- 1.96 x \frac{sd}{\sqrt(n)}$ [@orme2020, p. 94].

### Converting raw utilities

The validateHOT package also includes four functions designed to make the scores from both (A)CBC and MaxDiff analyses more interpretable, namely:

-   `att_imp()`: Converts the raw utilities of either an ACBC or CBC
    into importance scores for each attribute [see @orme2020, pp. 79-81]

-   `prob_scores()`: Converts the raw utilities of a MaxDiff to choice
    probabilities by applying the following procedures:

-   For unanchored MaxDiff: First, the scores are zero-centered,
    and then they are transformed by the following formula
    $\frac{exp^{U_i}}{(exp^{U_i} + a - 1)}$ [@chrzan2019, p. 56], where
    $U_i$ is the raw utility of item *i* and `a` is the number of items
    shown simultaneously per MaxDiff task.

-   For anchored MaxDiff a slightly different formula is applied:
    $\frac{exp^{U_i}}{(exp^{U_i} + a - 1)} * 100 * \frac{1}{a}$
    [@chrzan2019, p. 59].

-   `zc_diffs()`: Rescales the raw logit utilities to make them
    comparable across participants [@sawtooth2024, p. 330].

-   `zero_anchored()`: Rescales the raw logits of a MaxDiff to
    zero-centered diffs [@chrzan2019, p. 64].

### Data Frames provided by the validateHOT package

The package includes five data sets that help to better explain the
functions as well as the structure of the input, especially for the
`create_hot()` function.

-   `acbc`: Example data set with raw utilities of an ACBC study
    conducted in Sawtooth Software [@sawtooth2024]. The price was linear-coded while the other
    attributes were coded as part-worths [@sablotny-wackershauser2024;
    @sawtooth2024].

-   `acbc_interpolate`: Example data set with raw utilities of an ACBC
    study conducted in Sawtooth Software [@sawtooth2024] Price was piecewise-coded, another
    attribute was linear-coded while the other attributes were coded as
    part-worths [@sablotny-wackershauser2024; @sawtooth2024].

-   `cbc`: Example data set with raw utilities of a CBC study conducted
    in Sawtooth Software [@sawtooth2024] All attributes were coded as part-worth
    [@sablotny-wackershauser2024; @sawtooth2024].

-   `cbc_linear`: Example data set with raw utilities of a CBC study
    conducted in Sawtooth Software [@sawtooth2024] One attribute was linear-coded while the
    other attributes were part-worth coded [@sablotny-wackershauser2024;
    @sawtooth2024].

-   `maxdiff`: Example data set with raw utilities of a MaxDiff study
    conducted in Sawtooth Software [@sawtooth2024; @schramm2024].

## The story behind the validateHOT package

The validateHOT package was born out of teaching preference measurement seminars to students, many of whom have little to no prior experience with R. One of the chapters in this class is about model validation by
checking holdout tasks and we teach this, of course, in *R* 😍. We emphasize open science, and providing tools to run the analyses and share the code afterward. The validateHOT
package makes this process look easy 🤹‍♀️. Of course, there are other
great packages (i.e., Metricy package by @Metrics), however, these packages need some more data
wrangling to use the appropriate functions with the raw utilities, which
might be a burden or barrier some users.

Moreover, as @yang2018 report, commercial studies often do not use any
validation task. Again, the lack of experience in *R* could be one
explanation. Since these functions are not always implemented in other
software, this might be one reason why they do not include one simply
because they do not know how to apply it correctly. Having a package to
evaluate the validation task can be very beneficial from this
perspective.

## Installation

You can install the development version of the validateHOT package from
[GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("JoshSchramm94/validateHOT", dependencies = T, build_vignettes = T)
```

## Example

First, we load the package:

```{r}
library("validateHOT")
```

### Example I - CBC

Since *CBC's* are applied more commonly compared to *ACBC* and
*MaxDiff*, we will provide an example with a *CBC*. Let us begin with a
*CBC* where all of the attributes are part-worth coded and no interpolation was applied. Let us load the `cbc` data frame for this example
[@sablotny-wackershauser2024].

```{r}
data(cbc)
```

The data frame has a total of `r nrow(cbc)` participants and
`r ncol(cbc)` columns.

Now imagine you included a validation task with six alternatives plus a
no-buy alternative. We specify the `data` argument and the
`id`. Since we also have a *no-buy* alternative in our validation task,
we specify the `none` argument, otherwise we would have left it empty.
Afterwards, we define each alternative with the argument `prod.levels`.
If we look back at the data frame, we can see that the first alternative
in the holdout task (`c(3, 6, 10, 13, 16, 20, 24, 32, 35)`) is composed
of the following attribute levels `r colnames(cbc)[3]`,
`r colnames(cbc)[6]`, `r colnames(cbc)[10]`, `r colnames(cbc)[13]`,
`r colnames(cbc)[16]`, `r colnames(cbc)[20]`, `r colnames(cbc)[24]`,
`r colnames(cbc)[32]`, and `r colnames(cbc)[35]`.

As mentioned above, all the attributes are part-worth coded and the
alternatives have the same price as one of the levels shown (i.e., no
interpolation). Thus, we set `coding = c(rep(0, times = 9))`. Finally,
we specify the method, which is `method = "cbc"` in our case, and define
the column of the actual participant's choice (`choice`). If you run the
code, a data frame called `hot_cbc` will be returned to the global
environment.

> ❗ `create_hot()` takes both column names and column indexes. However,
> please be aware, if you include linear-coded or piecewise-coded, you
> **have** to provide the column indexes for the input in `prod.levels`.

```{r}
hot_cbc <- create_hot(
  data = cbc,
  id = "id",
  none = "none",
  prod.levels = list(
    c(3, 6, 10, 13, 16, 20, 24, 32, 35),
    c(3, 5, 10, 14, 16, 18, 22, 27, 35),
    c(4, 6, 9, 14, 15, 20, 25, 30, 36),
    c(4, 5, 10, 11, 16, 19, 26, 32, 34),
    c(2, 6, 8, 14, 16, 17, 26, 31, 36),
    c(2, 5, 7, 12, 16, 20, 26, 29, 33)
  ),
  coding = c(rep(0, times = 9)),
  method = "cbc",
  choice = "hot"
)
```

> In case you just need to create a market scenario, you can also leave
> the `choice` argument empty.

Sometimes iy you estimate a part-worth coded attribute but want to treat
this attribute as continuous in the validation task or market
simulations, please use the code `2` for this variable in the `coding`
argument.

Let us take a glimpse at the output, which shows the participants' total
raw utilities for each of the six alternatives that were included in the
validation task.

```{r}
head(hot_cbc)
```

In the next step, we would like to see how well our model (from which we
took the raw utilities) predicts the actual choices in the validation
task. First, we will run the `hitrate()` function. We specify the
`data`, the column names of the alternatives (`opts`; remember there are
six alternatives + the *no-buy* alternative), and finally the actual
choice (`choice`).

```{r}
hitrate(
  data = hot_cbc, # data frame
  opts = c(option_1:none), # column names of alternatives
  choice = choice # column name of choice
)
```

Next, we look at the magnitude of the mean absolute error by running the
`mae()` function. The arguments are the same as for the `hitrate()`
function.

```{r}
mae(
  data = hot_cbc, # data frame
  opts = c(option_1:none), # column names of alternatives
  choice = choice # column name of choice
)
```

Finally, let us test, how many participants would buy at least one of
three products, assuming that this is one potential assortment we would
like to offer to our consumers. We will use the `reach()` function. To
specify the bundles we are offering we use the `opts` argument in our
function.

```{r}
reach(
  data = hot_cbc, # data frame
  opts = c(option_1:option_3), # products that should be considered
  none = none # column name of none alternative
)
```

### Example II - CBC with linear-coded attribute(s)

In the second example, we again use a *CBC*, however, this time we show
how to use the package if one of the variables is linear-coded. All
other examples are provided in the accompanied vignette.

We are using the data frame `cbc_linear` [@sablotny-wackershauser2024].
Again, we first load the data frame.

```{r}
data(cbc_linear)
```

Next, we create the validation task to evaluate it in the next step. We
use the same validation task as defined above (i.e., six alternatives
plus the *no-buy* alternative). The only difference to the previous
example is that the last attribute (`price`) was linear-coded and
this time, we want to interpolate values.

Again, we first define data, namely the `id` as well as the `none`
alternative. Next, we define the `prod.levels` for each alternative.
Since we have one linear coded attribute, we need to specify the column
indexes instead of the column names in `prod.levels`. We tell
`create_hot()` that the last attribute needs to be interpolated by
specifying the `coding` argument accordingly. This tells us that the
first eight attributes are part-worth coded (`0`) while the last
attribute is linear-coded (`1`).

To interpolate the value, we need to provide `create_hot()` the
`interpolate.levels`. These **need** to be the same levels as provided
to Sawtooth Software [@sawtooth2024] or `ChoiceModelR`.
Extrapolation is allowed, however, `create_hot()` will give a warning in
case extrapolation is applied.

Next, we define the column of the linear coded variable (`lin.p`).
Again, we are running a CBC specified by the `method` argument. This
time, we would like to keep some of the variables in the data frame,
which we specify by using the `varskeep` argument. We only keep one
further variable, however, you can specify as many as you want. This
could be relevant if you would like to display results per group.
Finally, we define the actual choice (`choice`) in the validation task
and we are all set.

```{r}
hot_cbc_linear <- create_hot(
  data = cbc_linear,
  id = "id",
  none = "none",
  prod.levels = list(
    c(3, 6, 10, 13, 16, 20, 24, 32, 248.55),
    c(3, 5, 10, 14, 16, 18, 22, 27, 237.39),
    c(4, 6, 9, 14, 15, 20, 25, 30, 273.15),
    c(4, 5, 10, 11, 16, 19, 26, 32, 213.55),
    c(2, 6, 8, 14, 16, 17, 26, 31, 266.10),
    c(2, 5, 7, 12, 16, 20, 26, 29, 184.50)
  ),
  coding = c(rep(0, times = 8), 1),
  lin.p = 33,
  interpolate.levels = list(c(seq(from = 175.99, to = 350.99, by = 35))),
  method = "cbc",
  choice = "hot",
  varskeep = "group"
)
```

The next steps are the same as above. However, let us take a look at
some examples in which we display the results per group. Let us again
begin with the `hitrate()` function. To do so, we specify the column
name of the grouping variable in the `group` argument.

```{r}
hitrate(
  data = hot_cbc_linear, # data frame
  opts = c(option_1:none), # column names of alternatives
  choice = choice, # column name of choice
  group = group # column name of Grouping variable
)
```

Lastly, this time we also want to use a rescaling function, namely
`att_imp()` which tells us the importance of each attribute
included [@orme2020]. We need the data set with the raw logit
coefficients (`cbc_linear`; @sablotny-wackershauser2024). Next, we
define the `attrib` argument. Here, we need to specify each attribute
level for the corresponding level. Afterwards, we specify the coding
again, and since we have one linear coded attribute, we need to define
the `interpolate.levels` argument again, as we did for the
`create_hot()` function above. Finally, we set `res` to `agg`, which
tells `att_imp()` to display the aggregated results (to get results for each individual set `res` to `ind`).

```{r}
att_imp(
  data = cbc_linear,
  attrib = list(
    c(paste0("att1_lev", c(1:3))),
    c(paste0("att2_lev", c(1:2))),
    c(paste0("att3_lev", c(1:4))),
    c(paste0("att4_lev", c(1:4))),
    c(paste0("att5_lev", c(1:2))),
    c(paste0("att6_lev", c(1:4))),
    c(paste0("att7_lev", c(1:6))),
    c(paste0("att8_lev", c(1:6))),
    "price"
  ),
  coding = c(rep(0, times = 8), 1),
  interpolate.levels = list(c(seq(from = 175.99, to = 350.99, by = 35))),
  res = "agg"
)
```

For more examples, please see the accompanied vignette.

## References
